---
title: "Arimax/Time Series Forecasts"
author: Christian Url
date: 'Last Compiled `r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    
    code_folding: show
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
               collapse = FALSE,
               comment = "",
               strip.white = TRUE,
               warning = FALSE,
               message = FALSE,
               cache = TRUE,
               out.width = "100%",
               fig.align = "center")

```


## load Packages
```{r,  cache=FALSE}
library(data.table)
library(lubridate)
library(caret)
library(rai)
library(forecast)
library(parallel)
library(doParallel)
library(tidyverse)
```

## Paths
```{r}
datapath = "../data/"
res_path = paste0(datapath,"results/")
```


# Load Data {.tabset .tabset-fade}

## Volatility Index {-}
```{r}

vol_ind = read_delim(paste0(datapath,"country_volatility_index.csv")) %>%
  mutate(geo = str_split(Countries, " ", simplify = T)[,1], .before = 1,
         country = str_remove(Countries, ".*\\("), 
         country = str_extract(country, ".*(?=\\))"))

country_code_list = unlist(vol_ind$geo)
country_list = unlist(vol_ind$country)
country_code_list[country_list=="Greece"] <- "GR"
```

## Indicators {-}

```{r}
indicator = read_csv(paste0(datapath, "indicator.csv"))

ind_stats = indicator %>%
  select(country_code, mean, sd) %>%
  distinct()

indicator %>%
  group_by(country_code) %>%
  summarise(min_date = min(date), max_date = max(date)) %>%
  filter(max_date < as_date("2022-06-01"))
```
Check: ok! All time series range until either June or July.

## Trends {-}

# ARIMAX - Performance

To do - function: 

  * $n$ step ahead forecasts on rolling windows
  * select starting point
  
## Forecast functions {.tabset .tabset-fade}

### Arima {-}
```{r}
arima_models = function(input_ds, country_code_list, response_pos, n_ahead, ind_stats, transformation = "std", cl = detectCores()-1){
  
  registerDoParallel(cl)

    res_list = foreach(i = 1:length(country_code_list), .packages=c("tidyverse","forecast", "lubridate"), .multicombine = T, .combine = rbind) %dopar% {
    
      # Select country
      c_code = country_code_list[i]
      #ctry = country_list[i]
      
      ind_stats_act = ind_stats %>%
        filter(country_code == c_code)
      
      # Define time-series
      ts_def = input_ds %>%
        filter(country_code == c_code) %>%
        summarise(min_month = month(min(date)), 
                  min_year = year(min(date)))
    
      ind_ts = input_ds %>%
        filter(country_code == c_code) %>%
        .[,response_pos] %>%
        ts(frequency = 12, start = c(ts_def$min_year, ts_def$min_month))
    
      mod = auto.arima(ind_ts, max.d = 4, max.D = 4, max.order = 24, max.p = 12, max.q = 12)
      pred = forecast(mod, h = n_ahead)
      
      if(transformation=="std"){
        yhat = pred$mean * ind_stats_act$sd + ind_stats_act$mean
      }
      if(transformation=="log"){
        yhat = exp(pred$mean)
      }
      if(transformation=="sqrt"){
        yhat = pred$mean^2
      }
      
      coefs = as.data.frame(coef(mod)) %>% 
        rownames_to_column(var = "coef") %>%
        rename(val = `coef(mod)`) %>%
        pivot_wider(names_from = coef, values_from = val) %>%
        add_column(country_code = c_code, .before = 1) %>%
        add_column(sigma = mod$sigma2)
      
      arima_order = as.data.frame(arimaorder(mod)) %>% 
        rownames_to_column(var = "coef") %>%
        rename(val = `arimaorder(mod)`) %>%
        pivot_wider(names_from = coef, values_from = val) %>%
        add_column(country_code = c_code, .before = 1)
        
      
      yhat_list = list(tibble(date = zoo::as.Date(pred$mean),country_code = c_code, yhat = yhat, h=1:n_ahead, bic = mod$bic, 
                              aicc = mod$aicc, rmse = forecast::accuracy(mod)[2], mae = forecast::accuracy(mod)[3], 
                              acf = forecast::accuracy(mod)[7]),
                      coefs,
                      arima_order)
  }
  stopImplicitCluster()
  res_df = do.call(bind_rows, res_list[,1])
  coefs_df = do.call(bind_rows, res_list[,2])
  arima_order_df = do.call(bind_rows, res_list[,3])
  
  out_list = list(res_df, coefs_df, arima_order_df)
}
```

### ARIMAX with Fourier regression {-}

```{r}
arimax_fourier_models = function(input_ds, country_code_list, response_pos, n_ahead, ind_stats, 
                                 transformation = "std", cl = detectCores()-1, select_fourier = "bic"){
  
  registerDoParallel(cl)

    res_list = foreach(i = 1:length(country_code_list), .packages=c("tidyverse","forecast", "lubridate"), 
                       .multicombine = T, .combine = rbind) %dopar% {
    
      # Select country
      c_code = country_code_list[i]
      #ctry = country_list[i]
      
      ind_stats_act = ind_stats %>%
        filter(country_code == c_code)
      
      # Define time-series
      ts_def = input_ds %>%
        filter(country_code == c_code) %>%
        summarise(min_month = month(min(date)), 
                  min_year = year(min(date)))
    
      ind_ts = input_ds %>%
        filter(country_code == c_code) %>%
        .[,response_pos] %>%
        ts(frequency = 12, start = c(ts_def$min_year, ts_def$min_month))
    
      # Fourier models
      mod_out = list()
      stats_list = list()
      
      for (j in seq(6)) {
        mod_out[[j]] <- auto.arima(ind_ts, xreg = fourier(ind_ts, K = j), seasonal = FALSE, lambda = 0)
        stats_list[[j]] = tibble(bic=mod_out[[j]]$bic, aicc=mod_out[[j]]$aicc, sigma2=mod_out[[j]]$sigma2, loglik=mod_out[[j]]$loglik)
      }
      stats = do.call(rbind, stats_list)
      acc = do.call(rbind, map(mod_out, forecast::accuracy))
      
      # Determine best Fourier mod
      if(select_fourier == "aicc"){
        f_ord = which(stats$aicc == min(stats$aicc))
      }
      if(select_fourier == "aic"){
        f_ord = which(stats$aic == min(stats$aic))
      }
      if(select_fourier == "bic"){
        f_ord = which(stats$bic == min(stats$bic))
      }
      
      mod = mod_out[[f_ord]]
      pred = forecast(mod_out[[f_ord]], xreg=fourier(ind_ts, K=f_ord, h=n_ahead))
      
      if(transformation=="std"){
        yhat = pred$mean * ind_stats_act$sd + ind_stats_act$mean
      }
      if(transformation=="log"){
        yhat = exp(pred$mean)
      }
      if(transformation=="sqrt"){
        yhat = pred$mean^2
      }
      
      coefs = as.data.frame(coef(mod)) %>% 
        rownames_to_column(var = "coef") %>%
        rename(val = `coef(mod)`) %>%
        pivot_wider(names_from = coef, values_from = val) %>%
        add_column(country_code = c_code, .before = 1) %>%
        add_column(sigma = mod$sigma2)
      
      arima_order = as.data.frame(arimaorder(mod)) %>% 
        rownames_to_column(var = "coef") %>%
        rename(val = `arimaorder(mod)`) %>%
        pivot_wider(names_from = coef, values_from = val) %>%
        add_column(country_code = c_code, .before = 1)
        
      
      yhat_list = list(tibble(date = zoo::as.Date(pred$mean),country_code = c_code, yhat = yhat, h=1:n_ahead, bic = mod$bic, 
                              aicc = mod$aicc, rmse = forecast::accuracy(mod)[2], mae = forecast::accuracy(mod)[3], 
                              acf = forecast::accuracy(mod)[7]),
                      coefs,
                      arima_order)
  }
  stopImplicitCluster()
  res_df = do.call(bind_rows, res_list[,1])
  coefs_df = do.call(bind_rows, res_list[,2])
  arima_order_df = do.call(bind_rows, res_list[,3])
  
  out_list = list(res_df, coefs_df, arima_order_df)
}
```

### Rolling window wrapper {-}

```{r}
rolling_forecast = function(rolling_start = "2022-01-01", window_size = 2, input_ds = indicator, country_code_list, 
                           response_pos, ind_stats, transformation = "std", cl = detectCores()-1, 
                           model = "arima", select_fourier = "bic"){
  
  
  #Split data
  train = input_ds %>%
    filter(date < rolling_start)
  
  test = input_ds %>%
    filter(date >= rolling_start)


  ## for different window sizes - depricated
  # t1 = test %>%
  #   group_by(country_code) %>%
  #   summarise(nrow = n(),
  #             start_date = min(date),
  #             end_date = max(date)) %>%
  #   ungroup()
  # 
  # t2 = t1 %>%
  #   group_by(nrow) %>%
  #   group_map(~.x)
  # 
  # t3 = t2 %>% map(~left_join(.,t1))
  
  ## use only one window-size: 
  windows = test %>%
    group_by(country_code) %>%
    summarise(nrow = n(),
              start_date = min(date),
              end_date = max(date)) %>%
    ungroup() %>%
    filter(nrow == min(nrow)) %>%
    pull(nrow) %>%
    unique()
  
  train_control = train %>%
    group_by(country_code) %>%
    summarise(end_date = max(date))
  
  # create data set
  test_control = test %>%
    group_by(country_code) %>%
    summarise(nrow = n(),
              start_date = min(date),
              end_date = min(date) + months(1))  

  ########### RUNS ############
  out = list()
  
  runs = windows - window_size + 1
  
  for (i in seq(runs)){
    
    print(paste0("Run: ",i, "/",runs))
    
    train_filtered = train_control %>%
      inner_join(input_ds) %>%
      filter(date <= end_date) %>%
      select(!end_date)
  
    test_filtered = test_control %>%
      inner_join(test) %>%
      filter(date <= end_date & date >= start_date) %>%
      select(-c(nrow, end_date, start_date))
    
    
    if(model=="arima"){
      res = arima_models(input_ds = train_filtered, country_code_list = country_code_list, ind_stats = ind_stats , 
                         response_pos = response_pos, n_ahead = window_size, transformation = transformation)
    }
    if(model=="arimax_fourier"){
      res = arimax_fourier_models(input_ds = train_filtered, country_code_list = country_code_list, ind_stats = ind_stats , 
                         response_pos = response_pos, n_ahead = window_size, transformation = transformation,
                         select_fourier = select_fourier)
    }
    y_hat = do.call(bind_rows, res[[1]])

   stats = test_filtered %>%
    select(date, country_code, resid_total) %>%
    left_join(y_hat) %>%
    group_by(country_code) %>%
    summarise(rmse = caret::RMSE(yhat, resid_total),
              mae = caret::MAE(yhat, resid_total),
              window = i)
    
    out$results[[i]] = stats
    out$trace[[i]] = res
    
    train_control = train_control %>%
      mutate(end_date = end_date + months(1))
    
    test_control = test_control %>%
      mutate(start_date = start_date + months(1),
             end_date = end_date + months(1))
    
  }
  
  result = list()
  result$results = do.call(rbind,out$results)
  result$trace = out$trace
  
  return(result)
}



```

## Rolling window forecasts {.tabset .tabset-fade}

### Arima std {-}
```{r arima_roll}
arima_std = rolling_forecast(rolling_start = "2021-12-01", window_size = 2, input_ds = indicator, 
                             country_code_list = country_code_list, response_pos = 6, ind_stats = ind_stats, 
                             transformation = "std", model = "arima")

arima_res = arima_std$results %>%
  group_by(country_code) %>%
  summarise(rmse_mean = mean(rmse),
            mae_mean = mean(mae))

arima_res
```

Plots:

```{r}
preds = list()

for (i in 1:6){
  preds[[i]] = arima_std$trace[[i]][[1]]
}
preds_out = do.call(rbind,preds)

h1 = preds_out %>%
  arrange(country_code, date) %>%
  filter(h == 1) %>%
  inner_join(indicator)

h2 = preds_out %>%
  arrange(country_code, date) %>%
  filter(h == 2) %>%
  inner_join(indicator)

h1 %>%
  ggplot(mapping = aes(x = date, group = country_code)) +
  geom_line(aes(y = resid_total,color="Value")) + 
  geom_line(aes(y = yhat, color = "Pred")) + 
  facet_wrap(~country_code, ncol = 3, scales = "free_y")

h2 %>%
  ggplot(mapping = aes(x = date, group = country_code)) +
  geom_line(aes(y = resid_total,color="Value")) + 
  geom_line(aes(y = yhat, color = "Pred")) + 
  facet_wrap(~country_code, ncol = 3, scales = "free_y")
```


### Arima sqrt {-}
```{r arima_roll 2}
arima_sqrt = rolling_forecast(rolling_start = "2021-12-01", window_size = 2, input_ds = indicator, 
                             country_code_list = country_code_list, response_pos = 4, ind_stats = ind_stats, 
                             transformation = "sqrt", model = "arima")

arima_res2 = arima_sqrt$results %>%
  group_by(country_code) %>%
  summarise(rmse_mean = mean(rmse),
            mae_mean = mean(mae))

arima_res2
```

Plots:

```{r}
preds = list()

for (i in 1:6){
  preds[[i]] = arima_std$trace[[i]][[1]]
}
preds_out = do.call(rbind,preds)

h1 = preds_out %>%
  arrange(country_code, date) %>%
  filter(h == 1) %>%
  inner_join(indicator)

h2 = preds_out %>%
  arrange(country_code, date) %>%
  filter(h == 2) %>%
  inner_join(indicator)

h1 %>%
  ggplot(mapping = aes(x = date, group = country_code)) +
  geom_line(aes(y = resid_total,color="Value")) + 
  geom_line(aes(y = yhat, color = "Pred")) + 
  facet_wrap(~country_code, ncol = 3, scales = "free_y")

h2 %>%
  ggplot(mapping = aes(x = date, group = country_code)) +
  geom_line(aes(y = resid_total,color="Value")) + 
  geom_line(aes(y = yhat, color = "Pred")) + 
  facet_wrap(~country_code, ncol = 3, scales = "free_y")
```


### Arimax with fourier regressor {-}

```{r fourier_roll}
arimax_fourier = rolling_forecast(rolling_start = "2021-12-01", window_size = 2, input_ds = indicator, 
                             country_code_list = country_code_list, response_pos = 5, ind_stats = ind_stats, 
                             transformation = "log", model = "arimax_fourier")
```

```{r}
arima_fourier_res = arimax_fourier$results %>%
  group_by(country_code) %>%
  summarise(rmse_mean = mean(rmse),
            mae_mean = mean(mae))

arima_fourier_res
```


Plots:

```{r}
preds = list()

for (i in 1:6){
  preds[[i]] = arimax_fourier$trace[[i]][[1]]
}
preds_out = do.call(rbind,preds) 

id = tibble(id = rep(1:2,144))

h1 = preds_out %>%
  arrange(country_code, date) %>%
  filter(h == 1) %>%
  inner_join(indicator)

h2 = preds_out %>%
  arrange(country_code, date) %>%
  filter(h == 2) %>%
  inner_join(indicator)

h1 %>%
  ggplot(mapping = aes(x = date, group = country_code)) +
  geom_line(aes(y = resid_total,color="Value")) + 
  geom_line(aes(y = yhat, color = "Pred")) + 
  facet_wrap(~country_code, ncol = 3, scales = "free_y")

h2 %>%
  ggplot(mapping = aes(x = date, group = country_code)) +
  geom_line(aes(y = resid_total,color="Value")) + 
  geom_line(aes(y = yhat, color = "Pred")) + 
  facet_wrap(~country_code, ncol = 3, scales = "free_y")
```

### Comparing all tables {-}
```{r}
a = list()
a[[1]] = arima_res %>%
  add_column(transform = "std")
a[[2]] = arima_res2 %>%
  add_column(transform = "sqrt")
a[[3]] = arima_fourier_res %>%
  add_column(transform = "fourier")

overview = do.call(bind_rows, a) %>%
  arrange(country_code)

print(overview, n = nrow(overview))
```

### Select minimum {-}
```{r}
selector = overview %>%
  group_by(country_code) %>%
  filter(transform != "fourier") %>%
  summarise(rmse_min = min(rmse_mean),rmse_mean, transform) %>%
  filter(rmse_min==rmse_mean)

print(selector, n = nrow(selector))
write_csv(selector, file = paste0(res_path,"arima_selector.csv"))
```


# Forecasts {.tabset .tabset-fade}
## Arima {-}
```{r}
arima_out = list()

arima_out[[1]] = arima_models(input_ds = indicator,country_code_list = country_code_list, n_ahead=4, response_pos = 4, ind_stats = ind_stats, transformation = "sqrt")

arima_out[[2]] = arima_models(input_ds = indicator,country_code_list = country_code_list, n_ahead=4, response_pos = 6, ind_stats = ind_stats, transformation = "std")

write_csv(arima_out[[1]][[1]], file = paste0(res_path,"arima_sqrt.csv"))
write_csv(arima_out[[2]][[1]], file = paste0(res_path,"arima_std.csv"))
```

## Arimax Fourier {-}

```{r}
arimaxf_out = arimax_fourier_models(input_ds = indicator,country_code_list = country_code_list, n_ahead=4, response_pos = 5, ind_stats = ind_stats, transformation = "log")

write_csv(arimaxf_out[[1]], file = paste0(res_path,"arimax_fourier.csv"))
```

