---
title: "PPI indicator data"
author: Christian Url
date: 'Last Compiled `r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    
    code_folding: show
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
               collapse = FALSE,
               comment = "",
               strip.white = TRUE,
               warning = FALSE,
               message = FALSE,
               cache = FALSE,
               out.width = "100%",
               fig.align = "center")

```


## load Packages
```{r}
library(data.table)
library(lubridate)
library(rai)
library(caret)
library(forecast)
library(parallel)
library(doParallel)
library(tidyverse)
```

## Set paths
```{r}
datapath = "../data/"
ind_path = paste0(datapath, "indicators/")
```

## Functions
```{r}
scale2 <- function(x, na.rm = TRUE) (x - mean(x, na.rm = na.rm)) / sd(x, na.rm)
```


# Load Data

## Volatility Index
```{r}
vol_ind = read_delim(paste0(datapath,"country_volatility_index.csv")) %>%
  mutate(geo = str_split(Countries, " ", simplify = T)[,1], .before = 1)
```

## Indicators

Read indicator data and filter for valid countries. Countries are valid if they appear in the country volatility index.

```{r}
list.files(ind_path)
ind_in = fread(paste0(ind_path,"ppi.csv"))
ind_in[,1:7] %>% distinct()
ind_t1 = ind_in[,-c(1:7)] %>% 
  as_tibble() %>%
  mutate(date = as_date(paste0(TIME_PERIOD,"-01")),.before=1) %>%
  filter(geo %in% vol_ind$geo)
```

# Data cleansing

Build dataset:
```{r}
df_ind = ind_t1 %>%
  .[,c(1,2,4)] %>%
  filter(date >= as_date("2020-01-01")) %>%
  group_by(geo) %>%
  mutate(obs_value_std = scale2(OBS_VALUE), 
         mean=mean(OBS_VALUE), 
         sd = sd(OBS_VALUE),
         year = year(date),
         month = month(date)) %>%
  ungroup()

ind_stats  = df_ind %>%
  select(geo, mean, sd) %>%
  distinct()

geo_minmax = df_ind %>%
  group_by(geo) %>%
  summarise(min_year = year(min(date)), 
            max_year = year(max(date)),
            min_month = month(min(date)),
            max_month = month(max(date)))

print(geo_minmax, n=nrow(geo_minmax))
```

Plot of 5 years for all countries:
```{r plt_share, cache = T}
df_ind %>%
  filter(year >= 2016) %>%
  ggplot(aes(x = month, y=obs_value_std,group = interaction(factor(year),geo), color = factor(year))) + 
  geom_line() + 
  facet_wrap(~geo)
```

# Time series modeling 

Using arima for predicting September 2022

```{r multicore}
cl = detectCores()-1
registerDoParallel(cl)

geo_filter = geo_minmax %>%
  filter(max_year == 2022) %>% 
  .$geo 

res_list = foreach(i = 1:length(geo_filter), .packages=c("tidyverse","forecast"), .multicombine = T, .combine = rbind) %dopar% {
  ind_stats_act = ind_stats %>%
    filter(geo == geo_filter[i])
  
  ts_def = geo_minmax %>%
    filter(geo == geo_filter[i])

  ind_ts = df_ind %>%
    filter(geo == geo_filter[i]) %>%
    select(obs_value_std) %>%
    ts(frequency = 12, start = c(ts_def$min_year, ts_def$min_month))

  mod = auto.arima(ind_ts, max.d = 4, max.D = 4, max.order = 24, max.p = 12, max.q = 12)
  pred = forecast(mod, h = 4)

  yhat = pred$mean * ind_stats_act$sd + ind_stats_act$mean
  coefs = as.data.frame(coef(mod)) %>% 
    rownames_to_column(var = "coef") %>%
    rename(val = `coef(mod)`) %>%
    pivot_wider(names_from = coef, values_from = val) %>%
    add_column(geo = geo_filter[i], .before = 1) %>%
    add_column(sigma = mod$sigma2)
  
  arima_order = as.data.frame(arimaorder(mod)) %>% 
    rownames_to_column(var = "coef") %>%
    rename(val = `arimaorder(mod)`) %>%
    pivot_wider(names_from = coef, values_from = val) %>%
    add_column(geo = geo_filter[i], .before = 1)
    
  
  yhat_list = list(tibble(date = zoo::as.Date(pred$mean),geo=geo_filter[i], yhat = yhat, bic = mod$bic, 
                   rmse = forecast::accuracy(mod)[2], mae = forecast::accuracy(mod)[3], acf = forecast::accuracy(mod)[7]),
                   coefs,
                   arima_order)
}
res_df = do.call(bind_rows, res_list[,1])
coefs_df = do.call(bind_rows, res_list[,2])
arima_order_df = do.call(bind_rows, res_list[,3])

stopImplicitCluster()

res_df

write_csv(res_df, file="arima_forecasts_short.csv")
write_csv(coefs_df, file="arima_forecasts_coefs_short.csv")
write_csv(arima_order_df, file="arima_forecasts_order_short.csv")
```

overview over last forecast date:

```{r}
overview = res_df %>%
  group_by(geo) %>%
  summarise(min_date = min(date), max_date = max(date), bic = min(bic), rmse = min(rmse), mae = min(mae), acf = min(acf))

print(overview, n = nrow(overview))

print(filter(overview, min_date >= as_date("2022-07-01")), n = nrow(filter(overview, min_date >= as_date("2022-07-01"))))
print(arima_order_df, n = nrow(arima_order_df))
coefs_df
```

coefficients: 
```{r, results='asis'}
print(xtable::xtable(coefs_df), type="html")
```


*October forecasts:* 
```{r, results='asis'}
sept = res_df %>%
  filter(date == as_date("2022-10-01"))

print(xtable::xtable(sept), type="html")
```

*November forecasts:* 
```{r, results='asis'}
sept = res_df %>%
  filter(date == as_date("2022-11-01"))

print(xtable::xtable(sept), type="html")
```