---
title: "Google trends data"
author: Christian Url
date: 'Last Compiled `r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    
    code_folding: show
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
               collapse = FALSE,
               comment = "",
               strip.white = TRUE,
               warning = FALSE,
               message = FALSE,
               cache = TRUE,
               out.width = "70%",
               fig.align = "center")

```


## load Packages
```{r}
library(data.table)
library(tidyverse)
library(lubridate)
library(gtrendsR)
```

## Selectors

```{r}
country_code = "DE"
country = "Germany"
```


# Download data

Set the data path:
```{r}
datapath = "../data/"
trends_path = paste0(datapath, "googleTrendsAuto/")
```

Using the Package *gtrendsR*. First, find interesting categories:

```{r categories}
data("categories")
cat1 = categories %>%  
  mutate(cat = str_to_lower(name)) %>% 
  filter(str_detect(cat, "travel|holiday|hotel|camping|appartment|booking|rental|hike")) %>%
  distinct() %>%
  filter(str_detect(cat, "dvd|holidays|entertainment|luggage|party",  negate = T))

cat1
```

Categories regarding restaurants etc.:
```{r}
cat2 = categories %>%  
  mutate(cat = str_to_lower(name)) %>% 
  filter(str_detect(cat, "restaurant|bar|gastro|food|hostility")) %>%
  distinct()
cat2
```


# Country-level {.tabset .tabset-fade}

## `r country` {-}

Either more than 1 keyword and only 1 geo or vice versa.
```{r it1}
test1 = gtrends(keyword = country, 
        geo=c("DE", "GB", country_code),
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 67,
        onlyInterest = F
        )

test1$interest_over_time %>% 
  as_tibble() %>%
  mutate(date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits, group = keyword)) +  
  geom_line() + 
  facet_grid(~keyword)
head(test1$related_queries)
```

## Cities in Italy {-}

```{r it2}

test1_int = gtrends(keyword = c("Rome", "Italy", "Milano", "Sicily"), 
        geo=c("DE"),
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 67,
        onlyInterest = T
        )

test1_int$interest_over_time %>% 
  as_tibble() %>%
  mutate(date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits, group = keyword)) +  
  geom_line() + 
  facet_grid(~keyword)
test1_int$related_queries
```

## `r country` in Travel world {-}

```{r it3}
test2 = gtrends(keyword = country, 
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 67,
        onlyInterest = F
        )

test2$interest_over_time %>% 
  as_tibble() %>%
  mutate(date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits, group = keyword)) +  
  geom_line(size=1) + 
  facet_grid(~keyword)
test2$related_queries
```

## Hotels {-}

```{r it4}
test3 = gtrends(keyword = country, 
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 179,
        onlyInterest = F
        )

test3$interest_over_time %>% 
  as_tibble() %>%
  mutate(date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits, group = keyword)) +  
  geom_line(size=1) + 
  facet_grid(~keyword)
test3$related_queries
```

## Car rental {-}

```{r it5}
test4 = gtrends(keyword = country, 
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 205,
        onlyInterest = F
        )

test4$interest_over_time %>% 
  as_tibble() %>%
  mutate(date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits, group = keyword)) +  
  geom_line(size=1) + 
  facet_grid(~keyword)

test4$related_queries
```

## Apartment {-}
```{r it6}
test5 = gtrends(keyword = country, 
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 378,
        onlyInterest = F
        )

test5$interest_over_time %>% 
  as_tibble() %>%
  mutate(date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits, group = keyword)) +  
  geom_line(size=1) + 
  facet_grid(~keyword)

test5$related_queries
```

## Hiking {-}

```{r it7}
test6 = gtrends(keyword = country, 
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 542,
        onlyInterest = F
        )

test6$interest_over_time %>% 
  as_tibble() %>%
  mutate(date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits, group = keyword)) +  
  geom_line(size=1) + 
  facet_grid(~keyword)

test6$related_queries
```

## Travel Guides {-}

```{r it8}
test7 = gtrends(keyword = country, 
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 1011,
        onlyInterest = F
        )

test7$interest_over_time %>% 
  as_tibble() %>%
  mutate(date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits, group = keyword)) +  
  geom_line(size=1) + 
  facet_grid(~keyword)

test7$related_queries
```

# Food {.tabset .tabset-fade}
Maybe include some searches like restaurants, hikes, beaches for real-time estimation. But this can also bear non-results.

## Search for keywords {-}
```{r keywords1}

rest1 = gtrends(keyword = c("Restaurant","Bar","Beach", "Club", "Pizzeria"), 
        geo = country_code,
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 0,
        onlyInterest = F
        )

rest1$interest_over_time %>% 
  as_tibble() %>%
  mutate(hits = case_when(hits == "<1" ~ 0.5,
                          TRUE ~ as.numeric(hits)),
         hits = as.numeric(hits),
    date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits, group = keyword)) +  
  geom_line(size=1) + 
  facet_grid(~keyword)

head(rest1$related_queries)
```

## Category Food&Drink {-}

```{r fandd}
rest2 = gtrends(keyword = , 
        geo = country_code,
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 71, #food & drink
        onlyInterest = F
        )

rest2$interest_over_time %>% 
  as_tibble() %>%
  mutate(hits = case_when(hits == "<1" ~ 0.5,
                          TRUE ~ as.numeric(hits)),
         hits = as.numeric(hits),
    date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits)) +  
  geom_line(size=1) 
rest2$related_queries
```

## Category Restaurants {-}

```{r rest}
rest3 = gtrends(keyword = , 
        geo = country_code,
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 276, #restaurants
        onlyInterest = F
        )

rest3$interest_over_time %>% 
  as_tibble() %>%
  mutate(hits = case_when(hits == "<1" ~ 0.5,
                          TRUE ~ as.numeric(hits)),
         hits = as.numeric(hits),
    date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits)) +  
  geom_line(size=1) 

rest3$related_queries
```


## Category Groceries {-}

```{r groc1}
rest4 = gtrends(keyword = , 
        geo = country_code,
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 121, #groceries
        onlyInterest = F
        )

rest4$interest_over_time %>% 
  as_tibble() %>%
  mutate(hits = case_when(hits == "<1" ~ 0.5,
                          TRUE ~ as.numeric(hits)),
         hits = as.numeric(hits),
    date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits)) +  
  geom_line(size=1) 

rest4$related_queries
```

## Category fast food {-}

```{r ffod}
rest5 = gtrends(keyword = , 
        geo = country_code,
        time="today+5-y",
        gprop = c("web", "news", "images", "froogle", "youtube"),
        category = 918, #fast  food
        onlyInterest = F
        )

rest5$interest_over_time %>% 
  as_tibble() %>%
  mutate(hits = case_when(hits == "<1" ~ 0.5,
                          TRUE ~ as.numeric(hits)),
         hits = as.numeric(hits),
    date = as_date(date)) %>%
  ggplot(mapping=aes(x = date, y = hits)) +  
  geom_line(size=1) 

rest5$related_queries
```



